---
title: "Intro to Dataframes"
author: "Benji Lamp"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic R Dataframe Tutorial

## 1. Working with Data Frames
### What is a Data Frame?
A data frame is a table or a two-dimensional array-like structure in R. Each column in a data frame can contain different types of data (numeric, character, factor, etc.).

### Creating a Data Frame
You can create a data frame using the `data.frame()` function:
```{r}
# Creating a simple data frame
df <- data.frame(
  Name = c("Alice", "Bob", "Charlie"), # Strings
  Age = c(23, 35, 28), # Integers
  Gender = c("F", "M", "M")
)
```

### Viewing the Data Frame
To view the entire data frame, simply type the name of the data frame:
```{r}
df
```

You can also use the `head()` function to view the first few rows:
```{r}
head(df)
```

Use the `str()` function to view how any data variable is structured:

```{r}
str(df)
```


### Accessing Columns in a Data Frame
You can access individual columns in a data frame using the `$` operator or double square brackets `[[ ]]`:
```{r}
# Accessing the 'Name' column
df$Name

# Accessing the 'Age' column using double brackets
df[["Age"]]
```

### Accessing Rows in a Data Frame
To access specific rows, use the row number inside square brackets `[ ]`:
```{r}
# Accessing the first row
df[1, ]

# Accessing the third row
df[3, ]
```

### Accessing Specific Elements
To access a specific element, use both row and column indices inside square brackets `[row, column]`:

```{r}
# Accessing the element in the first row, second column
df[1, 2]

# Accessing the element in the second row, 'Gender' column
df[2, "Gender"]
```

### Adding a New Column
You can add a new column to a data frame like this:
```{r}
# Adding a 'Height' column
df$Height <- c(5.5, 6.0, 5.8)
df
```

### Removing a Column
To remove a column, assign `NULL` to it:
```{r}
# Removing the 'Gender' column
df$Gender <- NULL
df
```

## 2. Basic Functions with Data Frames
### Summary Statistics
You can quickly get summary statistics for each column in a data frame using the `summary()`function:
```{r}
summary(df)
```

### Filtering Data
You can filter rows in a data frame based on certain conditions:
```{r}
# Filtering rows where Age is greater than 25
df[df$Age > 25, ]

# Filtering rows where Name is 'Alice'
df[df$Name == "Alice", ]
```

### Sorting Data
You can sort a data frame based on a specific column using the `order()` function:
```{r}
# Sorting by Age in ascending order
df[order(df$Age), ]

# Sorting by Age in descending order
df[order(-df$Age), ]
```

## 3. Conclusion
This tutorial introduced you to the basics of R programming and working with data frames. Practice creating your own data frames, accessing elements, and performing basic operations to become more comfortable with R!


