---
title: "Intro to Data Structures"
author: "Benji Lamp"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Data Structures in R

## 1. Overview
Data structures are fundamental building blocks in any programming language. In R, data structures are used to store and organize data in a way that makes it easy to access, manipulate, and analyze. This tutorial will introduce you to the basic data structures in R: vectors, matrices, lists, and data frames.

## 2. Vectors
### What is a Vector?
A vector is the simplest data structure in R. It is a sequence of elements of the same type, such as numbers, characters, or logical values (TRUE/FALSE).

### Creating a Vector
You can create a vector using the `c()` function:
```{r}
# Numeric vector
numeric_vector <- c(1, 2, 3, 4, 5)

# Character vector
char_vector <- c("apple", "banana", "peach")

# Logical vector
logical_vector <- c(TRUE, FALSE, TRUE)
```

### Accessing Vector Elements
You can access elements of a vector using square brackets `[ ]` and specifying the index:
```{r}
# Accessing the second element
numeric_vector[2]
```

## 3. Matrices
### What is a Matrix?
A matrix is a two-dimensional data structure in R where each element is of the same type. It is essentially a vector with dimensions (rows and columns).

### Creating a Matrix
You can create a matrix using the `matrix()` function:
```{r}
# Creating a 3x3 numeric matrix
matrix_example <- matrix(1:9, nrow = 3, ncol = 3)
```

```{r}
print(matrix_example)
```


### Accessing Matrix Elements
You can access elements of a matrix by specifying the row and column indices:
```{r}
# Accessing the element in the first row, second column
matrix_example[1, 2]

# Accessing the entire second row
matrix_example[2, ]
```

## 4. Lists
### What is a List?
A list is a versatile data structure that can store elements of different types, such as numbers, strings, vectors, and even other lists.

### Creating a List
You can create a list using the `list()` function:
```{r}
# Creating a list with different types of elements
list_example <- list(
  name = "Benji Lamp",
  age = 30,
  scores = c(85, 90, 88)
)

print(list_example)
```

### Accessing List Elements
You can access elements of a list using the `$` operator or double square brackets `[[ ]]`:
```{r}
# Accessing the 'name' element
list_example$name

# Accessing the 'scores' element using double brackets
list_example[["scores"]]
```

## 5. Data Frames
### What is a Data Frame?
A data frame is a two-dimensional table-like structure where each column can contain different types of data. It is one of the most commonly used data structures in R for data analysis.

### Creating a Data Frame
You can create a data frame using the `data.frame()` function:
```{r}
# Creating a data frame
df <- data.frame(
  Name = c("Alice", "Lewis", "Mike"),
  Age = c(23, 35, 28),
  Gender = c("F", "M", "M")
)

print(df)
```

### Accessing Data Frame Elements
You can access elements in a data frame by specifying the row and column indices or by using the `$` operator to access columns:
```{r}
# Accessing the 'Name' column
df$Name

# Accessing the element in the first row, second column
df[1, 2]
```

## 6. Conclusion
Understanding these basic data structures—vectors, matrices, lists, and data frames—is crucial for working effectively in R. Each structure serves different purposes, and knowing when and how to use them will help you manage and analyze your data more efficiently.

