---
title: "Untitled"
author: "Benji Lamp"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Tutorial: Editing an R Markdown File

## Introduction
R Markdown is a powerful tool for creating dynamic documents that integrate code, results, and narrative. This guide will help you get started with editing an R Markdown file.

## 1. Open the R Markdown File
- Open RStudio.
- Navigate to the **File** menu, click on **Open File...**, and select the `.Rmd` file you want to edit.

## 2. Understand the Structure
An R Markdown file consists of three main sections:
1. **YAML Header**: Contains metadata about the document, such as title, author, and output format. It is located at the top of the file, enclosed in `---`.

   ```yaml
   ---
   title: "My First R Markdown"
   author: "Your Name"
   date: "2024-08-23"
   output: html_document
   ---
   ```
2. **Markdown Content**: This is where you write text, add images, links, and format your document using Markdown syntax.
   ```markdown
   ## Introduction
   This is an example of a basic R Markdown document.
   ```
3. **Code Chunks**: These are blocks of code that run in R. Code chunks are enclosed in triple backticks and start with `{r}`. The shortcut for creating an R code chunk is `ctrl + alt + i`
   ```{r}
   summary(cars)
   ```

## 3. Edit the YAML Header
- Update the `title`, `author`, and `date` fields to reflect your document’s details.
- Choose the desired output format by changing `output: html_document` to another format like `pdf_document` or `word_document`.

## 4. Add and Format Content
   
- To create bullet points, use asterisks or hyphens:

```markdown
   - First point
   - Second point
```
   
- To create numbered lists, use numbers followed by periods:

```markdown
   1. First item
   2. Second item
```

- To add emphasis, use asterisks or underscores:

```markdown
   *italic* or _italic_
   **bold** or __bold__
```

## 5. Edit Code Chunks
- Modify existing code or add new chunks by typing your code between the backticks:
   ````{r}
   # This is an R code chunk
   x <- 1:10
   mean(x)



